---
title: "Assignment 3"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment3q.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
```

------------------------------------------------------------------------

## Problems

\

1.  Create the following data frame

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
df = data.frame(
  Rank = c(0, 1, 2, 3),
  Age = c(28, 34, 29, 42),
  Name = c("Tom", "Jack", "Steve", "Ricky")
)

df
```

2.  Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://bryantstats.github.io/math421/data/us-states.csv).

    ```{r}
    covid = read.csv("https://covid19.who.int/who-data/vaccination-data.csv")

    ```

3.  Show the names of the variables in the data

```{r}
df = read.csv("https://bryantstats.github.io/math421/data/us-states.csv")
names(df)
```

4.  How many columns and rows the data have?

```{r}
dim(df)
```

5.  How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?

```{r}
colSums(is.na(df))
sum(is.na(df))
```

6.  What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.

    ```{r}
    class(df$date)

    df$date = as.Date(df$date)

    class(df$date)
    ```

7.  Capitalize the names of all the variables

    ```{r}
    names(df) = toupper(names(df))
    names(df)
    ```

8.  Find the average number of cases per day. Find the maximum cases a day.

    ```{r}
    mean(df$CASES)
    max(df$CASES)
    ```

9.  How many states are there in the data?

    ```{r}
    length(unique(df$STATE))
    ```

10. Create a new variable `weekdays` to store the weekday for each rows.

    ```{r}
    df$weekdays = weekdays(df$DATE)
    head(df$weekdays)
    ```

11. Create the categorical variable `death2` variable taking the values as follows

-   `has_death` if there is a death that day

-   `no_death` if there is no death that day

    ```{r}
    df$death2 = ifelse(df$DEATHS > 0, "has_death", "no_death")
    table(df$death2)
    prop.table(table(df$death2))
    ```

Find the frequency and relative frequency of `no_death` and `has_death`.

12. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)

    ```{r}
    summary(df$DEATHS)
    ```

13. Create the categorical variable `death3` variable taking the values as follows

-   `low_death` if the number of deaths smaller than the 25 percentile (Q1)

-   `mid_death` if the number of deaths from Q1 to Q3

-   `high_death` if the number of deaths greater than Q3

    ```{r}
    df$death3 = ifelse(df$DEATHS < 838, "low_death",
                       ifelse(df$DEATHS > 12674, "high_death", "mid_death"))
    table(df$death3)
    ```

14. Find the average cases in Rhode Island in 2021

    ```{r}
    mean(df$CASES[df$STATE == "Rhode Island" & format(df$DATE, "%Y") == "2021"])
    ```

15. Find the median cases by weekdays in Rhode Island in 2021

    ```{r}
    df1 = df[(df$STATE == "Rhode Island") & 
             (df$DATE >= "2021-01-01") & 
             (df$DATE <= "2021-12-31"),]

    by(data = df1$CASES, INDICES = df1$weekdays, FUN = median)
    ```

16. Compare the median cases in Rhode Island in June, July, August and September in 2021.

    ```{r}
    df1$month = format(df1$DATE, "%b")

    df2 = df1[df1$month %in% c("Jun", "Jul", "Aug", "Sep"),]

    by(data = df2$CASES, INDICES = df2$month, FUN = median)
    ```

17. Find your own dataset, import it and implement the following functions on the data

-   head

-   str

-   names

-   mean, min, max

-   table

-   is.na

-   colSums

-   class

-   cor

-   by

-   ifelse

-   case_when

```{r}
titanic = read.csv("titanic.csv")

# head
head(titanic)

# str
str(titanic)

# names
names(titanic)

# mean, min, max
mean(titanic$Fare)
min(titanic$Fare)
max(titanic$Fare)

# table
table(titanic$Sex)

# is.na
is.na(titanic)

# colSums
colSums(is.na(titanic))

# class
class(titanic$Age)

# cor
cor(titanic$Age, titanic$Fare, use = "complete.obs")

# by
by(data = titanic$Fare, INDICES = titanic$Pclass, FUN = mean)

# ifelse
titanic$survival = ifelse(titanic$Survived == 1, "Survived", "Did_Not_Survive")
table(titanic$survival)

# case_when
titanic$age_group = case_when(
  titanic$Age < 18 ~ "Child",
  titanic$Age < 60 ~ "Adult",
  titanic$Age >= 60 ~ "Senior"
)
table(titanic$age_group)
```

18. In the dataset in #17, practice the follows. You can reuse the code of 17.

-   Create a categorical variable from a continuous variable

-   From one categorical variable, create a new categorical variable with fewer categories

```{r}
# continuous variable to categorical variable
table(titanic$age_group)

# categorical variable with fewer categories
titanic$class_group = ifelse(
  titanic$Pclass == 1,
  "First_Class",
  "Other_Class"
)
table(titanic$class_group)
```

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](https://bryantstats.github.io/math421/data/titanic.csv)
