---
title: "Assignment 7 - Visualization - Some aspects of ggplot2"
format: html
editor: visual
---

***How to do it?***:

-   Open the Rmarkdown file of this assignment ([link](assignment7.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

------------------------------------------------------------------------

Use the data of your own. Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis.

------------------------------------------------------------------------

1.  Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)

    ```{r}
    df <- read.csv("C:/Users/student/OneDrive - Bryant University/math421assigments/titanic.csv")

    library(tidyverse)

    df %>%
      ggplot(aes(x = Pclass)) +
      geom_bar() +
      facet_wrap(~Survived) +
      labs(
        title = "Passenger Class by Survival Status",
        x = "Passenger Class",
        y = "Number of Passengers",
        caption = "Titanic Dataset"
      )
    ```

    The graph shows that third-class passengers had the highest number of deaths, while first-class passengers had the highest number of survivors. This suggests that passenger class played a role in survival.

2.  Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)

    ```{r}
    df %>%
      ggplot(aes(x = Age)) +
      geom_histogram(binwidth = 10) +
      facet_wrap(~Sex + Survived) +
      labs(
        title = "Passenger Age by Sex and Survival Status",
        x = "Age",
        y = "Number of Passengers",
        caption = "Titanic Dataset"
      )
    ```

    The graph shows that more female passengers survived than died, while male passengers had significantly more deaths. Most of the male passengers who died were between the ages of 20 and 40.

3.  Make a plot using `position ='dodge'`.

    ```{r}
    df %>%
      ggplot(aes(x = Pclass, fill = factor(Survived))) +
      geom_bar(position = "dodge") +
      labs(
        title = "Survival by Passenger Class",
        x = "Passenger Class",
        y = "Number of Passengers",
        fill = "Survived",
        caption = "Titanic Dataset"
      )
    ```

    The graph shows that first-class passengers had more survivors than deaths, while third-class passengers had significantly more deaths than survivors. This suggests that passengers in higher classes had a better chance of surviving.

4.  Make a plot using `position ='fill'`.

    ```{r}
    df %>%
      ggplot(aes(x = Sex, fill = factor(Survived))) +
      geom_bar(position = "fill") +
      labs(
        title = "Survival Proportions by Sex",
        x = "Sex",
        y = "Proportion of Passengers",
        fill = "Survived",
        caption = "Titanic Dataset"
      )
    ```

    The graph shows that about 74% of female passengers survived, compared to only 19% of male passengers. This shows that women had a much higher survival rate than men.

5.  Make a plot using `geom_col`.

    ```{r}
    df %>%
      group_by(Pclass) %>%
      summarise(AvgFare = mean(Fare, na.rm = TRUE)) %>%
      ggplot(aes(x = factor(Pclass), y = AvgFare)) +
      geom_col() +
      labs(
        title = "Average Ticket Fare by Passenger Class",
        x = "Passenger Class",
        y = "Average Fare",
        caption = "Titanic Dataset"
      )
    ```

    The graph shows that first-class passengers paid the highest average ticket fare at around 84, compared to 21 for second class and 14 for third class. This shows a large difference in ticket prices between passenger classes.

6.  Change the theme of one of your plots.

    ```{r}
    df %>%
      group_by(Pclass) %>%
      summarise(AvgFare = mean(Fare, na.rm = TRUE)) %>%
      ggplot(aes(x = factor(Pclass), y = AvgFare)) +
      geom_col() +
      labs(
        title = "Average Ticket Fare by Passenger Class",
        x = "Passenger Class",
        y = "Average Fare",
        caption = "Titanic Dataset"
      ) +
      theme_minimal()
    ```

    I used the minimal theme to remove the gray background and make the graph easier to read. The data stayed the same, but the graph looks cleaner.

7.  Make a plot using combo filter + ggplot

    ```{r}
    df %>%
      filter(Pclass == 3) %>%
      ggplot(aes(x = Age, fill = factor(Survived))) +
      geom_histogram(binwidth = 10, position = "dodge") +
      labs(
        title = "Age Distribution of Third-Class Passengers",
        x = "Age",
        y = "Number of Passengers",
        fill = "Survived",
        caption = "Titanic Dataset"
      )
    ```

    The graph shows that most third-class passengers who died were between 15 and 35 years old. There were fewer survivors than deaths across most age groups.

8.  Make a plot using combo group_by + summarise + geom_col

    ```{r}
    df <- read.csv("C:/Users/student/OneDrive - Bryant University/math421assigments/titanic.csv")
    df %>%
      group_by(Sex) %>%
      summarise(AvgAge = mean(Age, na.rm = TRUE)) %>%
      ggplot(aes(x = Sex, y = AvgAge)) +
      geom_col() +
      labs(
        title = "Average Passenger Age by Sex",
        x = "Sex",
        y = "Average Age",
        caption = "Titanic Dataset"
      )

    ggsave("average_age.png", width = 8, height = 5)
    ```

    The graph shows that male passengers had an average age of about 31, while female passengers had an average age of about 28. This shows that male passengers were slightly older on average.

9.  Save a plot and put a link of the plot to your Github Webpage. To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
[Link to the plot](abc.png)
```

10. In this question, we will use the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

-   [Data](https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df <- read_csv('https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv')
```

`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))
```

Do the follows.

-   Use `geom_point` to compare the vaccines used by date in Argentina in 2022.

    ```{r}
    library(tidyverse)

    vaccines <- read_csv("https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv")

    vaccines %>%
      filter(location == "Argentina",
             date >= as.Date("2022-01-01"),
             date <= as.Date("2022-12-31")) %>%
      ggplot() +
      geom_point(mapping = aes(x = date,
                               y = total_vaccinations,
                               color = vaccine)) +
      labs(
        title = "COVID-19 Vaccinations in Argentina (2022)",
        x = "Date",
        y = "Total Vaccinations",
        color = "Vaccine",
        caption = "COVID-19 Vaccination Dataset"
      )
    ```

    The graph shows that Sinopharm and Oxford/AstraZeneca had the highest vaccination totals in Argentina during 2022. Pfizer/BioNTech and Moderna showed steady increases, while CanSino had the lowest total.

-   Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots.

    ```{r}
    vaccines %>%
      filter(location == "Argentina",
             date >= as.Date("2022-01-01"),
             date <= as.Date("2022-12-31")) %>%
      ggplot() +
      geom_line(mapping = aes(x = date,
                              y = total_vaccinations,
                              color = vaccine)) +
      labs(
        title = "COVID-19 Vaccination Trends in Argentina (2022)",
        x = "Date",
        y = "Total Vaccinations",
        color = "Vaccine",
        caption = "COVID-19 Vaccination Dataset"
      )
    ```

11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Notice that the above codes could be piped in one

```{r, eval=FALSE}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Do the follows.

-   Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia

    ```{r}
    vaccines %>%
      filter(location %in% c("Japan", "South Korea")) %>%
      group_by(date, location) %>%
      summarise(total_vac = sum(total_vaccinations, na.rm = TRUE)) %>%
      ggplot(aes(x = date, y = total_vac, color = location)) +
      geom_point() +
      labs(
        title = "COVID-19 Vaccinations: Japan vs South Korea",
        x = "Date",
        y = "Total Vaccinations",
        color = "Country",
        caption = "COVID-19 Vaccination Dataset"
      )
    ```

    The graph shows that Japan had significantly more total vaccinations than South Korea. Both countries experienced an increase over time, but Japan's vaccination totals grew much faster.

-   Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021

    ```{r}
    vaccines %>%
      filter(location %in% c("Argentina", "Chile"),
             date >= as.Date("2021-01-01"),
             date <= as.Date("2021-12-31")) %>%
      group_by(date, location) %>%
      summarise(total_vac = sum(total_vaccinations, na.rm = TRUE)) %>%
      ggplot(aes(x = date, y = total_vac, color = location)) +
      geom_point() +
      labs(
        title = "COVID-19 Vaccinations: Argentina vs Chile (2021)",
        x = "Date",
        y = "Total Vaccinations",
        color = "Country",
        caption = "COVID-19 Vaccination Dataset"
      )
    ```

The graph shows that Chile initially had more total vaccinations than Argentina, but Argentina passed Chile around July 2021. By the end of the year, Argentina had about 80 million total vaccinations compared to Chile's 33 million.
