
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 6: Visualization - Aesthetic Mapping"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment6.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

### Questions on Your Selected Data

Choose a dataset and produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. 

-------
```{r}
library(tidyverse)

titanic <- read.csv("titanic.csv")

head(titanic)
```


1. For one continuous variable: Fare or Age

- Density
```{r}
titanic %>% ggplot()+
  geom_density(mapping = aes(x = Fare))
```
The density plot shows that most passengers paid lower fares. A smaller number of passengers paid much higher fares.

- Histogram 
```{r}
titanic %>% ggplot()+
  geom_histogram(mapping = aes(x = Fare))
```
The histogram shows that most fares were on the lower end, while only a few passengers paid very high fares.

- Boxplot
```{r}
titanic %>% ggplot()+
  geom_boxplot(mapping = aes(x = Age))
```
The boxplot shows that most passengers were around 20 to 40 years old. There were also several older passengers that appeared as outliers.

-------

2. For one categorical variable: Sex

- Barplot
```{r}
titanic %>% ggplot()+
  geom_bar(mapping = aes(x = Sex))
```
The barplot shows that there were more male passengers than female passengers on the Titanic.

-------

3. For two continuous variables: Age and Fare

- Scatter Plot
```{r}
titanic %>% ggplot()+
  geom_point(mapping = aes(x = Age, y = Fare, color = Sex))
```
The scatter plot does not show a strong relationship between age and fare. Both male and female passengers appear across a wide range of fares.

- Line Plot
```{r}
titanic %>% ggplot()+
  geom_line(mapping = aes(x = Age, y = Fare, color = Sex))
```
The line plot shows a lot of variation in fares across different ages for both males and females. There is no clear pattern of fare increasing or decreasing with age.

- Smooth Plot
```{r}
titanic %>% ggplot()+
  geom_smooth(mapping = aes(x = Age, y = Fare, color = Sex))
```
The smooth plot shows how the relationship between age and fare differs slightly between male and female passengers. Overall, there is not a strong consistent relationship.

-------

4. For one continuous + one categorical variables: Age and Sex

- Density
```{r}
titanic %>% ggplot()+
  geom_density(mapping = aes(x = Age, color = Sex))
```
The density plot shows that the age distributions for male and female passengers were fairly similar. Both groups had the highest concentration of passengers in their 20s.

- BoxPlot
```{r}
titanic %>% ggplot()+
  geom_boxplot(mapping = aes(x = Sex, y = Age))
```
The boxplot shows that male and female passengers had similar age distributions. The median age for males was slightly higher than for females.

-------

5. For two categorical variables: barplot: Sex and Pclass
```{r}
titanic %>% ggplot()+
  geom_bar(mapping = aes(x = Sex, fill = factor(Pclass)))
```
The barplot shows the number of male and female passengers in each passenger class. Males made up a larger number of passengers across the different classes.


-------

6. A plot that contains three variables: Age, Fare and Pclass
```{r}
titanic %>% ggplot()+
  geom_point(mapping = aes(x = Age, y = Fare, color = factor(Pclass)))
```
The plot shows the relationship between age and fare across the three passenger classes. First-class passengers generally paid higher fares, while second- and third-class passengers tended to pay lower fares.

