
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.
```{r}
library(tidyverse)

adult <- read_csv("adult_census_missing.csv")

colSums(is.na(adult))
```



2. Using function `aggr` the VIM package to plot the number of NAs for each column. 
```{r}
library(VIM)

aggr(adult)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
table(adult$workclass, useNA = "ifany")
table(adult$occupation, useNA = "ifany")
table(adult$native.country, useNA = "ifany")
```


4. Replace all the forms of missing values found to NA
```{r}
adult <- adult %>%
  mutate(across(everything(), ~replace(., . %in% c("?", "Not Known", "Unknown"), NA)))
colSums(is.na(adult))
```


5. Replot the number of NAs for each column. 
```{r}
aggr(adult)
```


6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
adult_drop <- drop_na(adult)

nrow(adult_drop)
```


7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
adult_fill <- adult %>%
  fill(everything(), .direction = "downup")
colSums(is.na(adult_fill))
```


8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 
```{r}
adult_replace <- adult %>%
  mutate(
    across(
      where(is.numeric),
      ~replace_na(., median(., na.rm = TRUE))
    ),
    across(
      where(is.character),
      ~replace_na(., names(sort(table(.), decreasing = TRUE))[1])
    )
  )
```

```{r}
colSums(is.na(adult_replace))
```




